---
title: "General Social Survey"
author: "Gracie Telles"
date: "Fall 2023"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```







The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  

# Question 1

Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: There is no relationship between gun ownership and opinion on the death penalty.

$H_A$: There is a relationship between gun ownership and opinion on the death penalty

## Methods

The death penalty is a categorical variable, and gun ownership is a categorical variable, therfor we use the cat-cat helper file.

## Results

## Graphical Descriptive Results

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")
```

The bar graph indicates that gun owners are more likely than non-gun owners to favor the death penalty.

## Numerical Descriptive Results

```{r}
table1 <- xtabs(~OWNGUN + CAPPUN,data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

The table indicates that 76% of gun owners are in favor of the death penalty, as opposed to the non-gun owners who only favor the death penalty by 58%.


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value of chi-squared test is 0.02022, which is less than 0.05.The p-value of the fisher exact test is 0.01651, which is less than 0.05 So we reject the null hypothesis- there is a relationship between gun ownership and being in favor with the death penalty. Gun owners are 2.3 times more likely to favor capital punishment than those who do not own a gun.

#  Question 2

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is no difference between years of education and those who did or did not vote in 2012.

$H_A$: There is no difference between years of education and those who did or did not vote in 2012.

## Methods

Years of education is a numerical variable, people who voted  is a categorical variable. Therfor we will use the numerical-cat helper file.

## Results

## Graphical Descriptive Results

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

The plot indicates that the people who voted in 2012 have a higher level of years of education, as opposed to those who did not vote.

## Numerical Descriptive Results

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

The table indicates that the median for  people who voted in 2012 is 16, and the median for the people that didn't vote in 2012 was 13. This corroborates the previous plot that the majority of voters were educated.

## Inferential Results

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

The t-test p-value is 0.0000001248, which is less than 0.05. Therefor we reject the null hypothesis- there is a difference between years of education and those who did or did not vote in 2012. 

#  Question 3

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  

$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE

## Methods

The average score on a wordsum test is a numerical variable, and the level of education is a categorical variable. Therefor we use the numerical~cat helper file is the helper file. 

## Results

## Graphical Descriptive Results

```{r}
# Create a box plot
ggplot(data=GSSdata, aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot() +
  xlab("Category") +
  ylab("WORDSUM") +
  ggtitle("Boxplot of Response Variable by Category")
```

The boxplot indicates that the WORSUM scores of the people with higher education like a Bachelor's, graduate, and junior college degrees are higher.

## Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

The table indicates that for those who earned a graduate, bachelors, or junior college level of education degree scored higher on the WORDSUM test (9-10). Those who earned a high school or LT high school degree scored lower (7-8).

## Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```

The p-value is 0.00000025 which is lower than 0.05. therefor we reject the null hypothesis- there is a difference in average score of WORDSUM test and level of education earned. We can conclude that the higher the education level, the higher the wordsum score.

# Discussion and Conclusion

In question 1 we used a cat-cat helper file to prove there is a correlation between gun owners and those that favor the death penalty. For question 2 we used a numerical-cat helper file to verify that there is a difference between the years of education a person received and whether or not they voted. For question 3, we used a numerical-cat helper file to reject the null hypothesis - there is a direct correlation between having a degree and the worsum score - the higher the education, the higher the word score.

