---
title: "General Social Survey"
author: "Gracie Telles"
date: "Fall 2023"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```







The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  

# Question 1

Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: There is no relationship between gun ownership and opinion on the death penalty.

$H_A$: There is a relationship between gun ownership and opinion on the death penalty

## Methods

Both of the variables are categorical, so we would use the Cat~Cat file.

## Results

## Graphical Descriptive Results

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")
```

The graph shows that the people who own a gun are most likely to favor the death penalty.The people who do not own a gun are less likely to favor the death penalty.

## Numerical Descriptive Results

```{r}
table1 <- xtabs(~OWNGUN + CAPPUN,data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

The summary shows us that the people who own a gun are more likely to be on favor of the death penalty and the people who don't own a gun are less likely to favor the death penalty.


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value of chi-squared test is 0.02022, which is less than 0.05.So we would reject the null hypothesis.

The p-value of the fisher exact test is 0.01651, which is less than 0.05.We reject the null hypothesis. We can conclude that there is a relationship between gun relationship and opinion in capitol punishment. People who own a gun are 2.3 times more likely to favor capitol punishment than those who do not own a gun.

#  Question 2

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?



## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

