---
title: "General Social Survey"
author: "Gracie Telles"
date: "Fall 2023"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```







The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  

# Question 1

Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: There is no relationship between gun ownership and opinion on the death penalty.

$H_A$: There is a relationship between gun ownership and opinion on the death penalty

## Methods

Both of the variables are categorical, so we would use the Cat~Cat file.

## Results

## Graphical Descriptive Results

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")
```

The graph shows that the people who own a gun are most likely to favor the death penalty.The people who do not own a gun are less likely to favor the death penalty.

## Numerical Descriptive Results

```{r}
table1 <- xtabs(~OWNGUN + CAPPUN,data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

The summary shows us that the people who own a gun are more likely to be on favor of the death penalty and the people who don't own a gun are less likely to favor the death penalty.


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value of chi-squared test is 0.02022, which is less than 0.05.So we would reject the null hypothesis.

The p-value of the fisher exact test is 0.01651, which is less than 0.05.We reject the null hypothesis. We can conclude that there is a relationship between gun relationship and opinion in capitol punishment. People who own a gun are 2.3 times more likely to favor capitol punishment than those who do not own a gun.

#  Question 2

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is no difference between years of education and those who did or did not vote in 2012.

$H_A$: There is no difference between years of education and those who did or did not vote in 2012.

## Methods

The response variable would be EDUC,and it is numerical variable. The predictor variable would be VOTE12, and it is a categorical variable.We would use the numerical~cat helper file.

## Results


## Graphical Descriptive Results

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

The plot shows us that the people who voted in 2012 have a higher level of years of education, and the people that didn't vote in 2012 have a lower level of years of education.

## Numerical Descriptive Results

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

The table shows that the people who voted in 2012 have a higher level of years of education, and the people that didn't vote in 2012 have a lower level of years of education. The mean of level of years of education for the people who voted was about 15 and the mean for those who didn't vote was about 13.

## Inferential Results

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

The t-test p-value here would be 0.0000001248, which would be less than 0.05. As a result, we would then we reject the null hypothesis, which was that there is no difference between years of education and those who did or did not vote in 2012. We conclude that there is a difference between years of education and those who did or did not vote in 2012

#  Question 3

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  

$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE

## Methods

The response variable would be WORDSUM, and it would be a numerical variable. The predictor variable would be DEGREE, and it is a categorical variable.Numerical~cat file is the helper file we would use for this. 

## Results

## Graphical Descriptive Results

```{r}
# Create a box plot
ggplot(data=GSSdata, aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot() +
  xlab("Category") +
  ylab("WORDSUM") +
  ggtitle("Boxplot of Response Variable by Category")
```

The graph above shows us that the WORSUM scores of the people with the following levels of education: a Bachelor's degree, graduate, high school, and junior college dregree, and less than high school.

## Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

The data set shows us the median and mean scores of the WORDSUM test based on the level of education of a person has earned. From this data we can see that those who earned a graduate,bachelors, or junior college level of education degree scored higher on the WORDSUM test, they scored around 9-10.Those who earned a high school degree and LT high school degree ended up scoring lower on the WORDSUM test,scoring around 7-8.

## Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```

The p-value is 0.00000025. Since it is lower than 0.05, reject the null hypothesis, which was that there was no difference in average score of WORDSUM test and level of education earned. This means that we can conclude that there is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

